---
output:
  pdf_document: default
  html_document: default
---
1. 2 pts. Regarding reproducibility, what is the main point of writing your own functions
and iterations?
# Reduces human error such as minimizing the need for copy pasting. Some functions may not be useful, and may not mend well to a different project as you may think.
2. 2 pts. In your own words, describe how to write a function and a for loop in R and how
they work. Give me specifics like syntax, where to write code, and how the results are
returned
# Functions are written using function(). The code goes inside {} and you use return() to get the result. Example:
add_one <- function(x) {
  return(x + 1)
}

# A for loop repeats code. Here's a simple version:

for (i in 1:5) {
  print(i)
}
# You can also store or manipulate data within each iteration.
```{r}


cities = read.csv("Cities.csv")

haversine <- function(lat1, lon1, lat2, lon2) {
  rad.lat1 <- lat1 * pi/180
  rad.lon1 <- lon1 * pi/180
  rad.lat2 <- lat2 * pi/180
  rad.lon2 <- lon2 * pi/180
  
  delta_lat <- rad.lat2 - rad.lat1
  delta_lon <- rad.lon2 - rad.lon1
  
  a <- sin(delta_lat / 2)^2 + cos(rad.lat1) * cos(rad.lat2) * sin(delta_lon / 2)^2
  c <- 2 * asin(sqrt(a)) 
  
  earth_radius <- 6378137
  distance_km <- (earth_radius * c)/1000
  
  return(distance_km)
}

auburn <- cities[cities$city == "Auburn", ]
nyc <- cities[cities$city == "New York", ]

haversine(nyc$lat, nyc$lon, auburn$lat, auburn$lon)
# Should return ~1367.854
```


```{r}
distance_df <- NULL

for (i in 1:nrow(cities)) {
  if (cities$city[i] != "Auburn") {
    dist <- haversine(cities$lat[i], cities$lon[i], auburn$lat, auburn$lon)
    temp <- data.frame(City1 = cities$city[i], 
                       City2 = "Auburn", 
                       Distance_km = dist)
    distance_df <- rbind(distance_df, temp)
  }
}

distance_df

head(distance_df,6)

```

